A Pandemic-Scale Ancestral Recombination Graph for SARS-CoV-2

### 2025-11-09 latex/main.tex
duplicate PDF cross refs

\documentclass[12pt,letterpaper]{article}
\usepackage[a4paper, total={7in, 10in}]{geometry}
\renewcommand{\familydefault}{\sfdefault}
\usepackage{graphicx}
\usepackage{helvet}
\usepackage{authblk}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{placeins}
\usepackage{orcidlink}
\usepackage[super,comma,sort&compress]
   {natbib}\bibliographystyle{numbered}
\usepackage[right]{lineno} \linenumbers
\usepackage[normalem]{ulem}

@@

Here we present sc2ts, a new method to infer ARGs for SARS-CoV-2  
that sequentially adds samples over time in a similar manner 
to UShER, while automatically detecting and incorporating recombination
in real-time. Based on the freely-available Viridian dataset, a curated
% NOTE: Viridian doesn't provide alignments, we do that with MAFFT
collection of SARS-CoV-2 genomes purged of various systematic errors and artifacts~\cite{Hunt2024},
we infer an ARG with 2.48 million whole genome sequences.
We show that the non-recombinant portions of this ARG are congruent with the UShER phylogeny built 
from the same dataset as well as the Pango nomenclature system~\cite{Rambaut2020}.
The ARG also accurately captures evolutionary features such as mutational 
spectra~\cite{Bloom2023} and deletions\cite{Li2023indels}.
We show that sc2ts automatically and accurately detects details about
known recombinant lineages that were painstakingly collated through a community effort and provides novel insights into the relationships between these
recombinant lineages. We compare the recombination events detected by 
sc2ts with the state-of-the-art combination of UShER and 
RIPPLES~\cite{Turakhia2022}, showing that sc2ts is more sensitive,
detecting more of the well-characterized recombination events 
curated by the Pango lineage designation community.
The output of sc2ts---incorporating
genealogical relationships, point mutations, deletions, and 
recombination---provides a basis for the study of SARS-CoV-2
that systematically accounts for all of these evolutionary forces.
This ARG, coupled with the rich suite of supporting software
in the tskit~\cite{Ralph2020} and VCF-Zarr~\cite{Czech2025} ecosystems,
enables pandemic-scale analysis on a laptop. 
These resources provide a strong platform for future research,
helping us to learn from the vast troves of data collected during the
pandemic and to prepare for the next.

>>>>

\documentclass[12pt,letterpaper]{article}
\usepackage[a4paper, total={7in, 10in}]{geometry}
\renewcommand{\familydefault}{\sfdefault}
\usepackage{graphicx}
\usepackage{helvet}
\usepackage{authblk}
\usepackage[hypertexnames=false]{hyperref}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{placeins}
\usepackage{orcidlink}
\usepackage[super,comma,sort&compress]
   {natbib}\bibliographystyle{numbered}
\usepackage[right]{lineno} \linenumbers
\usepackage[normalem]{ulem}

@@

Here we present sc2ts, a new method to infer ARGs for SARS-CoV-2  
that sequentially adds samples over time in a similar manner 
to UShER, while automatically detecting and incorporating recombination
in real-time. Based on the freely-available Viridian dataset, a curated
% NOTE: Viridian doesn't provide alignments, we do that with MAFFT
collection of SARS-CoV-2 genomes purged of various systematic errors and artifacts~\cite{Hunt2024},
we infer an ARG with 2.48 million whole genome sequences.
We show that the non-recombinant portions of this ARG are congruent with the UShER phylogeny built 
from the same dataset as well as the Pango nomenclature system~\cite{Rambaut2020}.
The ARG also accurately captures evolutionary features such as mutational 
spectra~\cite{Bloom2023} and deletions\cite{Li2023indels}.
We show that sc2ts automatically and accurately detects details about
known recombinant lineages that were painstakingly collated through a community effort and provides novel insights into the relationships between these
recombinant lineages. We compare the recombination events detected by 
sc2ts with the state-of-the-art combination of UShER and 
RIPPLES~\cite{Turakhia2022}, showing that sc2ts is more sensitive,
detecting more of the well-characterised recombination events 
curated by the Pango lineage designation community.
The output of sc2ts---incorporating
genealogical relationships, point mutations, deletions, and 
recombination---provides a basis for the study of SARS-CoV-2
that systematically accounts for all of these evolutionary forces.
This ARG, coupled with the rich suite of supporting software
in the tskit~\cite{Ralph2020} and VCF-Zarr~\cite{Czech2025} ecosystems,
enables pandemic-scale analysis on a laptop. 
These resources provide a strong platform for future research,
helping us to learn from the vast troves of data collected during the
pandemic and to prepare for the next.
